### 2022-12-19 source/paper3.tex
reorganizatin for reproducibility

   \node (designPrior) [below right = 3.5em and 2em of oData]
   {\begin{tabular}{c} \textbf{Design prior for model parameter} \\
      {\scriptsize $f(\theta \given x_{o}, \text{external
      knowledge})$}\\
    \end{tabular}};

  \node[rectangle, draw, rounded corners = 0.2em] (external) [right = 17em of oData]
  {\begin{tabular}{c} \textbf{Initial prior for model parameter} \\
     {\scriptsize $f(\theta \given \text{external knowledge})$} \\
   \end{tabular}};

@@

 \node (rData) [below = 4.5em of designPrior]
 {\begin{tabular}{c} \textbf{Predictive distribution of replication data} \\
    {\scriptsize $f(x_{r} \given n_{r}, x_{o}, \text{external
    knowledge})$}
  \end{tabular}};

@@

\node (PRS) [below = 4.5em of rData]
{\begin{tabular}{c} \textbf{Probability of replication success} \\
   {\scriptsize $\Pr(X_r \in S \given n_{r}, x_{o}, \text{external
    knowledge})$}
 \end{tabular}};

@@

  \caption{Schematic illustration of Bayesian sample size determination for
    replication studies. The original and replication data are denoted by
    $x_{o}$ and $x_{r}$, respectively. Both are assumed to come from a
    distribution with density/probability mass function denoted by
    $f(x_{i} \given \theta)$ for $i \in \{o, r\}$. An initial prior with density
    function $f(\theta \given \text{external knowledge})$ is assigned to the
    model parameter $\theta$.}
\label{fig3:SSDschema}
\end{figure}

@@

As the observed original data $x_{o}$, the yet unobserved replication data
$X_{r}$ are assumed to come from a distribution depending on the parameter
$\theta$. The parameter $\theta$ thus provides a link between the two studies,
and the knowledge obtained from the original study can be used to make
predictions about the replication. The central quantity for doing so is the
so-called \emph{design prior} of the parameter $\theta$, which we write as the
posterior distribution of $\theta$ based on the original data and an
\emph{initial prior} for $\theta$
\begin{align}
  \label{eq3:dp}
  f(\theta \given x_o, \text{external knowledge}) =
  \frac{f(x_o \given \theta) \, f(\theta \given \text{external knowledge})}{f(x_o \given
  \text{external knowledge})}.
\end{align}
The initial prior of $\theta$ may depend on external knowledge (\eg{} data from
other studies) and it represents the uncertainty about $\theta$ before observing
the original data. We will discuss common types of external knowledge in the
replication setting in the next Section. %~\ref{sec3:SS}.
The design prior~\eqref{eq3:dp} hence represents the state of knowledge and
uncertainty about the parameter $\theta$ before the replication is conducted,
and, along with an assumed replication sample size $n_{r}$, it can be used to
compute a predictive distribution for the replication data
\begin{align}
  \label{eq3:ftr}
  f(x_r \given n_r, x_o, \text{external knowledge})
  &= \int%_\Theta
  % \underbrace{f(x_r \given n_{r}, \theta)}_{\text{likelihood}}
    f(x_r \given n_{r}, \theta)
    % \, \underbrace{f(\theta \given x_o, \text{external knowledge})}_{\text{design prior}}
    \, f(\theta \given x_o, \text{external knowledge})
    \, \text{d}\theta.
\end{align}

After completion of the replication, the observed data $x_r$ will be analyzed in
some way to quantify to what extent the original result could be replicated. The
analysis may involve the original data (for example, a meta-analysis of the two
data sets) or it may only use the replication data. Typically, there is a
\emph{success region} $S$ which implies that if the replication data fall within
it ($x_r \in S$), the replication is successful. The \emph{probability of
  replication success} can thus be computed by integrating the predictive
density~\eqref{eq3:ftr} over $S$. To ensure a sufficiently conclusive replication
design, the sample size $n_{r}$ is determined such that the probability of
replication success is at least as large as a desired target probability of
success, here and henceforth denoted by $1 - \beta$. The required sample size
$n_r^*$ is then the smallest sample size which leads to a probability of
replication success of at least $1 - \beta$, \ie{}
\begin{align}
  \label{eq3:nr}
  n_r^* = \inf \left\{n_r : \Pr(X_{r} \in S \given
  n_r, x_{o}, \text{external knowledge}) \geq 1 - \beta \right\}.
\end{align}

@@

\section{Sample size determination in the normal-normal hierarchical model}
\label{sec3:SS}
We will now illustrate the general methodology from the previous section in the
\emph{normal-normal hierarchical model} where predictive distributions and the
probability of replication success can often be expressed in closed-form,
permitting further insight. % To conduct SSD for replication studies
It is pragmatic to adopt a meta-analytic perspective and use only study level
summary statistics instead of the raw study data since the raw data from the
original study are not always available to the replicators. Typically, the
underlying parameter $\theta$ is a univariate effect size quantifying the effect
on the outcome variable (\eg{} a mean difference, a log odds ratio, or a log
hazard ratio). The original and replication study can then be summarized through
an effect estimate $\that$, possibly the maximum likelihood estimate, and a
corresponding standard error $\sigma$, \ie{} $x_{o} = \{\that_{o}, \sigma_{o}\}$
and $x_{r} = \{\that_{r}, \sigma_{r}\}$. Effect estimates and standard errors
are routinely reported in research articles or can, under some assumptions, be
computed from $p$-values and confidence intervals. As in the conventional
meta-analytic framework \citep{Sutton2001}, we further assume that for study
$k \in \{o, r\}$ the (suitably transformed) effect estimate $\hat{\theta}_k$ is
approximately normally distributed around a study specific effect size
$\theta_k$ and with (known) variance equal to its squared standard error
$\sigma_k^2$, here and henceforth denoted by
$\hat{\theta}_k \given \theta_k \sim \Nor(\theta_k, \sigma_k^2)$. The standard
error $\sigma_k$ is typically of the form $\sigma_k = \lambda/\sqrt{n_k}$ with
$\lambda^{2}$ some unit variance and $n_{k}$ the sample size. The ratio of the
original to the replication variance is thus the ratio of the replication to the
original sample size
\begin{align*}
  c = \sigma^2_o/\sigma^2_r = n_r/n_o,
\end{align*}
which is often the main focus of SSD as it quantifies how much the replication
sample $n_{r}$ size needs to be changed compared to the original sample size
$n_{o}$. Depending on the effect size type, this framework might require slight
modifications \citep[see \eg{}][ch. 2.4]{Spiegelhalter2004}.

@@

If heterogeneity is expected, there are different approaches for specifying
$\tau^{2}$. A domain expert may subjectively assess how much heterogeneity is to
be expected due to the change in laboratory, study population, and other
factors. An alternative is to take an estimate from the literature, \eg{} from
multisite replication projects or from systematic reviews. Finally, one can also
specify an upper limit of ``tolerable heterogeneity''. This approach is similar
to specifying a minimal clinically relevant difference in classical power
analysis in the sense that a true replication effect size which is intolerably
heterogeneous from the original effect size is not relevant to be detected. An
absolute \citep[ch. 5.7.3]{Spiegelhalter2004} and a relative approach
\citep{Held2020c} can be considered. In the absolute approach, a value of
$\tau^{2}$ is chosen such that a suitable range of study-specific effect sizes
is not larger than an effect size difference considered negligible. For example,
when 95\% of the study specific effect sizes should not vary more than a small
effect size \eg{} $d = 0.2$ on standardized mean difference
scale based on the \citet{Cohen1992} effect size classification, this would lead
to $\tau = d/(2 \cdot 1.96) \approx 0.05$. % Similarly, one
% could consider the distribution of the difference of two effect sizes
% $\theta_{1} - \theta_{2} \given \theta \sim \Nor(0, 2\tau^{2})$ and choose
% $\tau$ such that the, \eg{} the 95\% range is not larger than some desired
% value.
In the relative approach, $\tau^{2}$ is specified relative to the variance of
the original estimate $\sigma^{2}_{o}$ using field conventions for tolerable
relative heterogeneity. For example, in the Cochrane guidelines for systematic
reviews \citep{Deeks2019} a value of
$I^{2} = \tau^{2}/(\tau^{2} + \sigma^{2}_{o}) = 40\%$ is classified as
``negligible'', which translates to
$\tau^{2} = \sigma^{2}_{o}/(1/I^{2} - 1) = (2\sigma^{2}_{o})/3$.

@@

We now assume that the one-sided original $p$-value was significant at some
level $\alpha$, \ie{}
\mbox{$p_{o} = 1 - \Phi(\that_{o}/\sigma_{o}) \leq \alpha$}. Replication success
at level $\alpha$ % with the two-trials
is then achieved if the replication \mbox{$p$-value} is also significant, \ie{}
$p_{r} = 1 - \Phi(\that_{r}/\sigma_{r}) \leq \alpha$, which implies a success
region
\begin{align*}
  S_{\scriptscriptstyle \text{2TR}}
  = \left[\zalpha\, \sigma_{r}, \infty \right),
\end{align*}
where $\zalpha$ is the $1 - \alpha$ quantile of the standard
normal distribution. The probability of replication success is thus given by
\begin{align}
  \label{eq3:2TRpros}
  \Pr(\that_{r} \in S_{\scriptscriptstyle \text{2TR}} \given \that_{o}, \sigma_{o}, \sigma_{r})
  = % \Phi\left(\frac{\muthatr - \zalpha \, \sigma_{r}}{\sqrt{\sigma^{2}_{r} +
  % \tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right)
  \Phi\left(\frac{\muthatr - \zalpha \, \sigma_{r}}{\sigmathatr}\right)
\end{align}
with $\Phi(\cdot)$ the standard normal cumulative distribution function and
$\muthatr$ and $\sigmathatr$ the mean and standard deviation of the predictive
distribution~\eqref{eq3:fthetar}. %, the latter also depending on $\sigma_{r}$.
Importantly, by decreasing the standard error $\sigma_{r}$ (through increasing
the sample size $n_{r}$), the probability of replication
success~\eqref{eq3:2TRpros} cannot become arbitrarily large but is bounded from
above by
\begin{align}
  \label{eq3:limP2TR}
  \text{limPr}_{\scriptscriptstyle \text{2TR}} =
  \Phi\left(\frac{ \muthatr}{\sqrt{
\tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right).
\end{align}
The required replication standard error $\sigma_{r}^{*}$ to achieve a target
probability of replication success
$1 - \beta < \text{limPr}_{\scriptscriptstyle \text{2TR}}$ can now be obtained
by equating~\eqref{eq3:2TRpros} to $1 - \beta$ and solving for $\sigma_{r}$. This
leads to
\begin{align}
  \label{eq3:ssd2tr}
  \sigma_{r}^{*} =
  \frac{\muthatr \zalpha - \zbeta\sqrt{
  (\zalpha^{2} - \zbeta^{2})\left\{\tau^{2} + (\sigma^{2}_{o} +
  \tau^{2})/(1 + 1/g)\right\} + \muthatr^{2}}}{\zalpha^{2} - \zbeta^{2}}
\end{align}
for $\alpha < \beta$. The standard error $\sigma_{r}^{*}$ can subsequently be
translated in a sample size. The translation depends on the type of effect size,
for instance, for SMD effect sizes we can use the approximation
$n_{r}^{*} \approx 4/(\sigma^*_{r})^{2}$ from
earlier. %Section~\ref{sec3:example}.
Moreover, by assuming a standard error of the form
$\sigma_{r} = \lambda/\sqrt{n_{r}}$ and plugging in the parameters of the
``conditional'' design prior %from Section~\ref{sec3:designpredictive}
\mbox{($\tau^{2} = 0$, $\mutheta = \that_{o}$, $g \downarrow 0$)}, we obtain the
well-known sample size formula \citep[ch. 3.3]{Matthews2006}
\begin{align*}
  n_{r}^{*} = \frac{(\zalpha + \zbeta)^{2}}{(\that_{o}/\lambda)^{2}}
\end{align*}
for a one-sided significance test at level $\alpha$ with power $1 - \beta$ to
detect the original effect estimate $\that_{o}$. The formula~\eqref{eq3:ssd2tr}
thus generalizes standard sample size calculation to take into account the
uncertainty of the original estimate, between-study heterogeneity and other
types of external knowledge.

\subsubsection{Fixed effects meta-analysis}
The data from the original and replication studies are sometimes pooled via
fixed-effects meta-analysis. The pooled effect estimate $\that_{m}$ and standard
error $\sigma_{m}$ are then given by
\begin{align*}
  &\that_{m} =
    \left(\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r}\right)\sigma^{2}_{m}&
&\text{and}&                                                                                       &\sigma_{m} = \left(1/\sigma^{2}_{o} + 1/\sigma^{2}_{r}\right)^{-1/2},&
\end{align*}
and they are also equivalent to the mean and standard deviation of a posterior
distribution for the effect size $\theta$ based on the data from both studies
and an initial flat prior for $\theta$. The success
region %for the replication effect estimate $\that_{r}$
\begin{align}
  \label{eq3:SMA}
  S_{\scriptscriptstyle \text{MA}}
  = \left[\sigma_{r} \zalpha\sqrt{1 + \sigma^{2}_{r}/\sigma^{2}_{o}} -
  (\that_{o} \sigma^{2}_{r})/\sigma^{2}_{o},
  \infty \right)
\end{align}
then corresponds to both replication success defined via a one-sided
meta-analytic $p$-value being smaller than level $\alpha$, \ie{}
$p_{m} = 1 - \Phi(\that_{m}/\sigma_{m}) \leq \alpha$, or to replication success
defined via a Bayesian posterior probability
$\Pr(\theta > 0 \given \that_{o}, \that_{r}, \sigma_{o}, \sigma_{r}) \geq 1 - \alpha$.
Based on the success region~\eqref{eq3:SMA} and an assumed standard error
$\sigma_{r}$, the probability of replication success can be computed by
\begin{align}
  \label{eq3:MApros}
  \Pr(\that_{r} \in S_{\scriptscriptstyle \text{MA}} \given \that_{o}, \sigma_{o}, \sigma_{r})
  = % \Phi\left(\frac{\muthatr - \sigma_{r} \zalpha\sqrt{1 + \sigma^{2}_{r}/\sigma^{2}_{o}}
%   + (\that_{o} \sigma^{2}_{r})/\sigma^{2}_{o}}{\sqrt{\sigma^{2}_{r} +
%   \tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right).
  \Phi\left(\frac{\muthatr - \sigma_{r} \zalpha\sqrt{1 + \sigma^{2}_{r}/\sigma^{2}_{o}}
  + (\that_{o} \sigma^{2}_{r})/\sigma^{2}_{o}}{\sigmathatr}\right).
\end{align}
As for the two-trials rule, the probability~\eqref{eq3:MApros} cannot be made
arbitrarily large by decreasing the standard error $\sigma_{r}$ but is bounded
from above by $\text{limPr}_{\text{2TR}}$ defined in~\eqref{eq3:limP2TR}.
% However, as the original data already provide some information, the
% probability of replication success is also bounded from below.
The required standard error $\sigma_{r}^{*}$ to achieve a target probability of
replication success $1 - \beta < \text{limPr}_{\text{2TR}}$ can be computed
numerically using root finding algorithms.

>>>>

   \node (designPrior) [below right = 3.5em and 2em of oData]
   {\begin{tabular}{c} \textbf{Design prior for model parameter} \\
      {\scriptsize $f(\theta \given x_{o}, \mathrm{external~
      knowledge})$}\\
    \end{tabular}};

  \node[rectangle, draw, rounded corners = 0.2em] (external) [right = 17em of oData]
  {\begin{tabular}{c} \textbf{Initial prior for model parameter} \\
     {\scriptsize $f(\theta \given \mathrm{external~knowledge})$} \\
   \end{tabular}};

@@

 \node (rData) [below = 4.5em of designPrior]
 {\begin{tabular}{c} \textbf{Predictive distribution of replication data} \\
    {\scriptsize $f(x_{r} \given n_{r}, x_{o}, \mathrm{external~
    knowledge})$}
  \end{tabular}};

@@

\node (PRS) [below = 4.5em of rData]
{\begin{tabular}{c} \textbf{Probability of replication success} \\
   {\scriptsize $\Pr(X_r \in S \given n_{r}, x_{o}, \mathrm{external
    knowledge})$}
 \end{tabular}};

@@

  \caption{Schematic illustration of Bayesian sample size determination for
    replication studies. The original and replication data are denoted by
    $x_{o}$ and $x_{r}$, respectively. Both are assumed to come from a
    distribution with density/probability mass function denoted by
    $f(x_{i} \given \theta)$ for $i \in \{o, r\}$. An initial prior with density
    function $f(\theta \given \mathrm{external~knowledge})$ is assigned to the
    model parameter $\theta$.}
\label{fig3:SSDschema}
\end{figure}

@@

As the observed original data $x_{o}$, the yet unobserved replication data
$X_{r}$ are assumed to come from a distribution depending on the parameter
$\theta$. The parameter $\theta$ thus provides a link between the two studies,
and the knowledge obtained from the original study can be used to make
predictions about the replication. The central quantity for doing so is the
so-called \emph{design prior} of the parameter $\theta$, which we write as the
posterior distribution of $\theta$ based on the original data and an
\emph{initial prior} for $\theta$
\begin{align}
  \label{eq3:dp}
  f(\theta \given x_o, \mathrm{external~knowledge}) =
  \frac{f(x_o \given \theta) \, f(\theta \given \mathrm{external~knowledge})}{f(x_o \given
  \mathrm{external~knowledge})}.
\end{align}
The initial prior of $\theta$ may depend on external knowledge (\eg{} data from
other studies) and it represents the uncertainty about $\theta$ before observing
the original data. We will discuss common types of external knowledge in the
replication setting in the next Section. %~\ref{sec3:SS}.
The design prior~\eqref{eq3:dp} hence represents the state of knowledge and
uncertainty about the parameter $\theta$ before the replication is conducted,
and, along with an assumed replication sample size $n_{r}$, it can be used to
compute a predictive distribution for the replication data
\begin{align}
  \label{eq3:ftr}
  f(x_r \given n_r, x_o, \mathrm{external~knowledge})
  &= \int%_\Theta
  % \underbrace{f(x_r \given n_{r}, \theta)}_{\mathrm{likelihood}}
    f(x_r \given n_{r}, \theta)
    % \, \underbrace{f(\theta \given x_o, \mathrm{external~knowledge})}_{\mathrm{design prior}}
    \, f(\theta \given x_o, \mathrm{external~knowledge})
    \, \mathrm{d}\theta.
\end{align}

After completion of the replication, the observed data $x_r$ will be analyzed in
some way to quantify to what extent the original result could be replicated. The
analysis may involve the original data (for example, a meta-analysis of the two
data sets) or it may only use the replication data. Typically, there is a
\emph{success region} $S$ which implies that if the replication data fall within
it ($x_r \in S$), the replication is successful. The \emph{probability of
  replication success} can thus be computed by integrating the predictive
density~\eqref{eq3:ftr} over $S$. To ensure a sufficiently conclusive replication
design, the sample size $n_{r}$ is determined such that the probability of
replication success is at least as large as a desired target probability of
success, here and henceforth denoted by $1 - \beta$. The required sample size
$n_r^*$ is then the smallest sample size which leads to a probability of
replication success of at least $1 - \beta$, \ie{}
\begin{align}
  \label{eq3:nr}
  n_r^* = \inf \left\{n_r : \Pr(X_{r} \in S \given
  n_r, x_{o}, \mathrm{external~knowledge}) \geq 1 - \beta \right\}.
\end{align}

@@

\section{Sample size determination in the normal-normal hierarchical model}
\label{sec3:SS}
We will now illustrate the general methodology from the previous section in the
\emph{normal-normal hierarchical model} where predictive distributions and the
probability of replication success can often be expressed in closed-form,
permitting further insight. % To conduct SSD for replication studies
It is pragmatic to adopt a meta-analytic perspective and use only study level
summary statistics instead of the raw study data since the raw data from the
original study are not always available to the replicators. Typically, the
underlying parameter $\theta$ is a univariate effect size quantifying the effect
on the outcome variable (\eg{} a mean difference, a log odds ratio, or a log
hazard ratio). The original and replication study can then be summarized through
an effect estimate $\that$, possibly the maximum likelihood estimate, and a
corresponding standard error $\sigma$, \ie{} $x_{o} = \{\that_{o}, \sigma_{o}\}$
and $x_{r} = \{\that_{r}, \sigma_{r}\}$. Effect estimates and standard errors
are routinely reported in research articles or can, under some assumptions, be
computed from $p$-values and confidence intervals. As in the conventional
meta-analytic framework \citep{Sutton2001}, we further assume that for study
$k \in \{o, r\}$ the (suitably transformed) effect estimate $\hat{\theta}_k$ is
approximately normally distributed around a study specific effect size
$\theta_k$ and with (known) variance equal to its squared standard error
$\sigma_k^2$, here and henceforth denoted by
$\hat{\theta}_k \given \theta_k \sim \Nor(\theta_k, \sigma_k^2)$. The standard
error $\sigma_k$ is typically of the form $\sigma_k = \lambda/\sqrt{n_k}$ with
$\lambda^{2}$ some unit variance and $n_{k}$ the sample size. The ratio of the
original to the replication variance is thus the ratio of the replication to the
original sample size
\begin{align*}
  c = \sigma^2_o/\sigma^2_r = n_r/n_o,
\end{align*}
which is often the main focus of SSD as it quantifies how much the replication
sample $n_{r}$ size needs to be changed compared to the original sample size
$n_{o}$. Depending on the effect size type, this framework might require slight
modifications \citep[see \eg{}][chapter 2.4]{Spiegelhalter2004}.

@@

If heterogeneity is expected, there are different approaches for specifying
$\tau^{2}$. A domain expert may subjectively assess how much heterogeneity is to
be expected due to the change in laboratory, study population, and other
factors. An alternative is to take an estimate from the literature, \eg{} from
multisite replication projects or from systematic reviews. Finally, one can also
specify an upper limit of ``tolerable heterogeneity''. This approach is similar
to specifying a minimal clinically relevant difference in classical power
analysis in the sense that a true replication effect size which is intolerably
heterogeneous from the original effect size is not relevant to be detected. An
absolute \citep[chapter 5.7.3]{Spiegelhalter2004} and a relative approach
\citep{Held2020c} can be considered. In the absolute approach, a value of
$\tau^{2}$ is chosen such that a suitable range of study-specific effect sizes
is not larger than an effect size difference considered negligible. For example,
when 95\% of the study specific effect sizes should not vary more than a small
effect size \eg{} $d = 0.2$ on standardized mean difference
scale based on the \citet{Cohen1992} effect size classification, this would lead
to $\tau = d/(2 \cdot 1.96) \approx 0.05$. % Similarly, one
% could consider the distribution of the difference of two effect sizes
% $\theta_{1} - \theta_{2} \given \theta \sim \Nor(0, 2\tau^{2})$ and choose
% $\tau$ such that the, \eg{} the 95\% range is not larger than some desired
% value.
In the relative approach, $\tau^{2}$ is specified relative to the variance of
the original estimate $\sigma^{2}_{o}$ using field conventions for tolerable
relative heterogeneity. For example, in the Cochrane guidelines for systematic
reviews \citep{Deeks2019} a value of
$I^{2} = \tau^{2}/(\tau^{2} + \sigma^{2}_{o}) = 40\%$ is classified as
``negligible'', which translates to
$\tau^{2} = \sigma^{2}_{o}/(1/I^{2} - 1) = (2\sigma^{2}_{o})/3$.

@@

We now assume that the one-sided original $p$-value was significant at some
level $\alpha$, \ie{}
\mbox{$p_{o} = 1 - \Phi(\that_{o}/\sigma_{o}) \leq \alpha$}. Replication success
at level $\alpha$ % with the two-trials
is then achieved if the replication \mbox{$p$-value} is also significant, \ie{}
$p_{r} = 1 - \Phi(\that_{r}/\sigma_{r}) \leq \alpha$, which implies a success
region
\begin{align*}
  S_{\scriptscriptstyle \mathrm{2TR}}
  = \left[\zalpha\, \sigma_{r}, \infty \right),
\end{align*}
where $\zalpha$ is the $1 - \alpha$ quantile of the standard
normal distribution. The probability of replication success is thus given by
\begin{align}
  \label{eq3:2TRpros}
  \Pr(\that_{r} \in S_{\scriptscriptstyle \mathrm{2TR}} \given \that_{o}, \sigma_{o}, \sigma_{r})
  = % \Phi\left(\frac{\muthatr - \zalpha \, \sigma_{r}}{\sqrt{\sigma^{2}_{r} +
  % \tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right)
  \Phi\left(\frac{\muthatr - \zalpha \, \sigma_{r}}{\sigmathatr}\right)
\end{align}
with $\Phi(\cdot)$ the standard normal cumulative distribution function and
$\muthatr$ and $\sigmathatr$ the mean and standard deviation of the predictive
distribution~\eqref{eq3:fthetar}. %, the latter also depending on $\sigma_{r}$.
Importantly, by decreasing the standard error $\sigma_{r}$ (through increasing
the sample size $n_{r}$), the probability of replication
success~\eqref{eq3:2TRpros} cannot become arbitrarily large but is bounded from
above by
\begin{align}
  \label{eq3:limP2TR}
  \mathrm{limPr}_{\scriptscriptstyle \mathrm{2TR}} =
  \Phi\left(\frac{ \muthatr}{\sqrt{
\tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right).
\end{align}
The required replication standard error $\sigma_{r}^{*}$ to achieve a target
probability of replication success
$1 - \beta < \mathrm{limPr}_{\scriptscriptstyle \mathrm{2TR}}$ can now be obtained
by equating~\eqref{eq3:2TRpros} to $1 - \beta$ and solving for $\sigma_{r}$. This
leads to
\begin{align}
  \label{eq3:ssd2tr}
  \sigma_{r}^{*} =
  \frac{\muthatr \zalpha - \zbeta\sqrt{
  (\zalpha^{2} - \zbeta^{2})\left\{\tau^{2} + (\sigma^{2}_{o} +
  \tau^{2})/(1 + 1/g)\right\} + \muthatr^{2}}}{\zalpha^{2} - \zbeta^{2}}
\end{align}
for $\alpha < \beta$. The standard error $\sigma_{r}^{*}$ can subsequently be
translated in a sample size. The translation depends on the type of effect size,
for instance, for SMD effect sizes we can use the approximation
$n_{r}^{*} \approx 4/(\sigma^*_{r})^{2}$ from
earlier. %Section~\ref{sec3:example}.
Moreover, by assuming a standard error of the form
$\sigma_{r} = \lambda/\sqrt{n_{r}}$ and plugging in the parameters of the
``conditional'' design prior %from Section~\ref{sec3:designpredictive}
\mbox{($\tau^{2} = 0$, $\mutheta = \that_{o}$, $g \downarrow 0$)}, we obtain the
well-known sample size formula \citep[chapter 3.3]{Matthews2006}
\begin{align*}
  n_{r}^{*} = \frac{(\zalpha + \zbeta)^{2}}{(\that_{o}/\lambda)^{2}}
\end{align*}
for a one-sided significance test at level $\alpha$ with power $1 - \beta$ to
detect the original effect estimate $\that_{o}$. The formula~\eqref{eq3:ssd2tr}
thus generalizes standard sample size calculation to take into account the
uncertainty of the original estimate, between-study heterogeneity and other
types of external knowledge.

\subsubsection{Fixed effects meta-analysis}
The data from the original and replication studies are sometimes pooled via
fixed-effects meta-analysis. The pooled effect estimate $\that_{m}$ and standard
error $\sigma_{m}$ are then given by
\begin{align*}
  &\that_{m} =
    \left(\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r}\right)\sigma^{2}_{m}&
&\mathrm{and}&                                                                                       &\sigma_{m} = \left(1/\sigma^{2}_{o} + 1/\sigma^{2}_{r}\right)^{-1/2},&
\end{align*}
and they are also equivalent to the mean and standard deviation of a posterior
distribution for the effect size $\theta$ based on the data from both studies
and an initial flat prior for $\theta$. The success
region %for the replication effect estimate $\that_{r}$
\begin{align}
  \label{eq3:SMA}
  S_{\scriptscriptstyle \mathrm{MA}}
  = \left[\sigma_{r} \zalpha\sqrt{1 + \sigma^{2}_{r}/\sigma^{2}_{o}} -
  (\that_{o} \sigma^{2}_{r})/\sigma^{2}_{o},
  \infty \right)
\end{align}
then corresponds to both replication success defined via a one-sided
meta-analytic $p$-value being smaller than level $\alpha$, \ie{}
$p_{m} = 1 - \Phi(\that_{m}/\sigma_{m}) \leq \alpha$, or to replication success
defined via a Bayesian posterior probability
$\Pr(\theta > 0 \given \that_{o}, \that_{r}, \sigma_{o}, \sigma_{r}) \geq 1 - \alpha$.
Based on the success region~\eqref{eq3:SMA} and an assumed standard error
$\sigma_{r}$, the probability of replication success can be computed by
\begin{align}
  \label{eq3:MApros}
  \Pr(\that_{r} \in S_{\scriptscriptstyle \mathrm{MA}} \given \that_{o}, \sigma_{o}, \sigma_{r})
  = % \Phi\left(\frac{\muthatr - \sigma_{r} \zalpha\sqrt{1 + \sigma^{2}_{r}/\sigma^{2}_{o}}
%   + (\that_{o} \sigma^{2}_{r})/\sigma^{2}_{o}}{\sqrt{\sigma^{2}_{r} +
%   \tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right).
  \Phi\left(\frac{\muthatr - \sigma_{r} \zalpha\sqrt{1 + \sigma^{2}_{r}/\sigma^{2}_{o}}
  + (\that_{o} \sigma^{2}_{r})/\sigma^{2}_{o}}{\sigmathatr}\right).
\end{align}
As for the two-trials rule, the probability~\eqref{eq3:MApros} cannot be made
arbitrarily large by decreasing the standard error $\sigma_{r}$ but is bounded
from above by $\mathrm{limPr}_{\mathrm{2TR}}$ defined in~\eqref{eq3:limP2TR}.
% However, as the original data already provide some information, the
% probability of replication success is also bounded from below.
The required standard error $\sigma_{r}^{*}$ to achieve a target probability of
replication success $1 - \beta < \mathrm{limPr}_{\mathrm{2TR}}$ can be computed
numerically using root finding algorithms.

\subsubsection{Effect size equivalence test}
% Both the two-trials rule and the fixed effects meta-analysis approach define
% replication success via measures of evidence against the null hypothesis
% $H_{0} \colon \theta = 0$. Other approaches define replication success via
% compatibility of the effect estimates from original and replication study. For
% instance,
\citet{Anderson2016} proposed a method for quantifying replicability based on
effect size equivalence. Under normality, replication success at level $\alpha$
is achieved if the $(1 - \alpha)$ confidence interval for the effect size
difference $\theta_{r} - \theta_{o}$
\begin{align*}
  \that_{r} - \that_{o} \pm \zalphatwo \sqrt{\sigma^{2}_{r} + \sigma^{2}_{o}}
\end{align*}
is fully inside an equivalence region $[-\Delta, \Delta]$ defined via the
pre-specified margin $\Delta > 0$. This procedure corresponds to rejecting the
null hypothesis $H_{0} \colon |\theta_{r} - \theta_{o}| > \Delta$ in an
equivalence test, and it implies a success region for the replication effect
estimate $\that_{r}$ given by
\begin{align}
  \label{eq3:equivalence}
  S_{\scriptscriptstyle \mathrm{E}}
  = \left[\that_{o} - \Delta + \zalphatwo \sqrt{\sigma^{2}_{o} +
  \sigma^{2}_{r}}, \that_{o} + \Delta - \zalphatwo
  \sqrt{\sigma^{2}_{o} + \sigma^{2}_{r}}\right]
\end{align}
for $\Delta \geq \zalphatwo \sqrt{\sigma_{o}^{2} + \sigma^{2}_{r}}$. For too
small margins ($\Delta < \zalphatwo \sqrt{\sigma_{o}^{2} + \sigma^{2}_{r}}$),
the success region~\eqref{eq3:equivalence} becomes the empty set meaning that
replication success is impossible.
% Typically the original study is already conducted but the standard error of the
% replication study can still be changed
% Whether or
% not replication success is achievable depends on the standard error of the
% original estimate $\sigma_{o}$, and in particular when the margin is chosen too
% small ($\Delta \leq \zalphatwo \sigma_{o}$), replication success is impossible
% regardless of how small the replication standard error $\sigma_{r}$ will be.
Assuming now that the margin is large enough,
% ($\Delta \geq \zalphatwo \sqrt{\sigma_{o}^{2} + \sigma^{2}_{r}}$),
the probability
of replication success can be computed by
\begin{align}
  \label{eq3:porsEqu}
  \Pr(\that_{r} \in S_{\scriptscriptstyle \mathrm{E}} \given \that_{o}, \sigma_{o}, \sigma_{r})
  &= % \Phi\left(\frac{\that_{o} + \Delta - \zalphatwo
%   \sqrt{\sigma^{2}_{o} + \sigma^{2}_{r}} -\muthatr}{\sqrt{\sigma^{2}_{r} +
%   \tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right) \nonumber \\
% &\phantom{=} -
%   \Phi\left(\frac{\that_{o} - \Delta + \zalphatwo
%   \sqrt{\sigma^{2}_{o} + \sigma^{2}_{r}} -\muthatr}{\sqrt{\sigma^{2}_{r} +
%   \tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right).
    \Phi\left(\frac{\that_{o} + \Delta - \zalphatwo
  \sqrt{\sigma^{2}_{o} + \sigma^{2}_{r}} -\muthatr}{\sigmathatr}\right) \nonumber \\
&\phantom{=} -
  \Phi\left(\frac{\that_{o} - \Delta + \zalphatwo
  \sqrt{\sigma^{2}_{o} + \sigma^{2}_{r}} -\muthatr}{\sigmathatr}\right).
\end{align}
As with the previous methods, the probability~\eqref{eq3:porsEqu} cannot be made
arbitrarily large by decreasing the replication standard error $\sigma_{r}$, but
is bounded by
\begin{align*}
  \mathrm{limPr}_{\scriptscriptstyle \mathrm{E}} =
  \Phi\left(\frac{\that_{o} + \Delta - \zalphatwo\sigma_{o}
  -\muthatr}{\sqrt{\tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right)
  -\Phi\left(\frac{\that_{o} - \Delta + \zalphatwo\sigma_{o}
  -\muthatr}{\sqrt{\tau^{2} + (\sigma^{2}_{o} + \tau^{2})/(1 + 1/g)}}\right).
\end{align*}
The required replication standard error $\sigma_{r}^{*}$ to achieve a target
probability of replication success
$1 - \beta < \mathrm{limPr}_{\scriptscriptstyle \mathrm{E}}$ can again be computed
numerically.
